---
title: "EDA_HONEST_aim_1"
author: "Sarah Holzer"
date: "`r Sys.Date()`"
output:
  html_document: 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load, echo=FALSE, include=FALSE}
# Loading packages
library(tidyverse)
library(lme4)
library(lmerTest)
library(rptR)
library(Matrix)
library(ggplot2)
library(ggdist)
library(dplyr)
library(scales)
library(naniar)
library(table1)

# Load aggregated datasets honest and honest_DEF
load("/Volumes/research$/HONEST Team/Data/aim_1_processed_data/07_FINAL_datasets.RData")

# honest contains all observations without counting for inclusion and exclusion criteria at the hospital level while honest_DEF is the final dataset the inclusion/exclusion criteria were applied.

```

```{r dataset preparation for EDA, echo=FALSE, include=FALSE}
# Removing variables not needed for EDA
honest_DEF <- honest_DEF [, -which(names(honest_DEF) %in% c("LOHN_ARZT", "LOHN_PFLEGE", "pct.male_nurse", "pct.male_physician", "pct.notCH_nurse", "pct.notCH_physician", "PFLEGETAGE_STATIONAER", "AUSTRITTE_STATIONAER"))]

# Relocating equipment
honest_DEF <- honest_DEF %>% relocate (equipment, .before = AMBULANTE_KONSULTATIONEN)

# Inserting monthly wage/FTE variable in honest_DEF and reposition it near the original vars
honest_DEF$month_wageFTE_nurses <- honest_DEF$wageFTE_nurses/13
honest_DEF <- honest_DEF %>% relocate (month_wageFTE_nurses, .after = wageFTE_nurses)

honest_DEF$month_wageFTE_physicians <- honest_DEF$wageFTE_physicians/13
honest_DEF <- honest_DEF %>% relocate (month_wageFTE_physicians, .after = wageFTE_physicians)

# Inserting log(wage/FTE) variable in honest_DEF and reposition it near the original vars
honest_DEF$log_wageFTE_nurses <- log(honest_DEF$wageFTE_nurses)
honest_DEF <- honest_DEF %>% relocate (log_wageFTE_nurses, .after = wageFTE_nurses)

honest_DEF$log_wageFTE_physicians <- log(honest_DEF$wageFTE_physicians)
honest_DEF <- honest_DEF %>% relocate (log_wageFTE_physicians, .after = wageFTE_physicians)

# Inserting age^2 variable in honest_DEF and reposition it near the original vars
honest_DEF$age2.mean_nurse <- (honest_DEF$age.mean_nurse)^2
honest_DEF <- honest_DEF %>% relocate (age2.mean_nurse, .after = age.mean_nurse)

honest_DEF$age2.mean_physician <- (honest_DEF$age.mean_physician)^2
honest_DEF <- honest_DEF %>% relocate (age2.mean_physician, .after = age.mean_physician)

```





## 1. About the project

### Project Title
***NURSES’ AND PHYSICIANS’ WAGES DRIVERS IN SWISS HOSPITALS: A RETROSPECTIVE OBSERVATIONAL TIME SERIES CROSS SECTIONAL STUDY***

### Description
In this retrospective observational time series cross-sectional study based on de-identified data regularly collected by the Federal Statistical Office (BFS) which comprises information on Swiss acute care and specialized hospitals, as well as on staff (including nurses and physicians) employed in these hospitals from 2014 to 2020. After data cleaning and preparation, we conducted descriptive analysis. Eventually, we ran different regression models to determine the factors affecting wages for nurses and physicians.


## 2. Data Sources, Acquisition, and Cleaning

### Data Sources 
The original data sets are the typo_ks_personal_2014-2020, typo_ks_leistungen_2014-2020, typo_ks_kostentraeger_2014-2020, typo_ks_fibu_2014-2020, typo_ks_allg_2014-2020 and Leistung count in the KS_Typol_2014_2020 folder. This folder is located in the HONEST Team folder which is part of the Research$ folder on the nurseserve.urz.p.unibas.ch server.

Further information about the dataset can be found here: https://www.bfs.admin.ch/bfs/de/home/statistiken/gesundheit/erhebungen/ks.html 

### Data Acquisition
The FSO provided de-identified data on hospitals and specialized care clinics in Switzerland from 2014 to 2020 based on a data protection contract. According to this contract (chapter 4.2), the panel data of the same statistics, i.e. the corresponding data from the survey dates 2014 - 2020, may be linked with each other. 

Since the sample consisted of routinely collected, and de-identified data, it was exempt from ethics approval.

#### Inclusion/Exclusion criteria for final dataset `honest_DEF`:

- Inclusion criteria for acute care facilities: Swiss acute care hospitals and specialized acute care clinics with a physician full-time equivalent (FTE) of at least 1% of the overall FTEs of the hospital.
- Personal inclusion criteria: Full- and part-time working nurses and physicians, including those in training, aged over 18 and under 65.
- Employment contract requirement: Only nurses and physicians with a contract with an included hospital at the end of the year (31.12.) were included to avoid double-counting of staff in training or those who changed jobs.

## 3. Exploratory Data Analysis

### Univariate Analysis

In the following section you will see the distribution of each variable. Including summary information and a density plot for each variable.



```{r EDA, echo=FALSE, include=TRUE, warning=FALSE, message=FALSE}
# explore_complete_cases <- function(honest_DEF) {
#   # Subset the honest_DEF to include only complete cases
#   complete_data <- honest_DEF[complete.cases(honest_DEF), ]
#   
#   # Loop through all numeric variables in the complete honest_DEF
#   for (col in names(complete_data)) {
#     if (is.numeric(complete_data[[col]])) {
#       cat("Summary for", col, ":\n")
#       print(summary(complete_data[[col]]))
#       
#       # cat("\nHistogram for", col, ":\n")
#       # hist(complete_data[[col]], main = col)
#       
#       cat("\nDensity Plot for", col, ":\n")
#       plot(density(complete_data[[col]]), main = col)
#     }
#   }
# }
# 
# # Example usage:
# explore_complete_cases(honest_DEF)


# ----

# Creating vector with my variables
numerical_vars <- sapply(honest_DEF, is.numeric)
numerical_vars <- names(numerical_vars)[numerical_vars]

# Loop through numerical variables and create rain cloud plots
for (var in numerical_vars) {
  plot_title <- paste("Rain Cloud Plot for", var)
  
  # Create individual plot for each numerical variable
  plot <- ggplot(honest_DEF, aes(x = "", y = !!sym(var))) + 
    ggdist::stat_halfeye( # make sure that this is visualizing the raw data
      adjust = 0.5, 
      width = 0.6, 
      .width = 0, 
      justification = -0.3, 
      point_colour = NA
    ) + 
    geom_boxplot(
      width = 0.25, 
      outlier.shape = NA
    ) +
    geom_point(
      aes(y = !!sym(var)),
      size = 1.5,
      alpha = 0.2,
      position = position_jitter(  # CAVE jittering 
        seed = 1, width = 0.1
      )
    ) +
    labs(title = plot_title)
  
  # Print or save the individual plot
  print(plot)
}


```




### Variables for nurses regression 

```{r scatterplots variables, echo=FALSE, warning=FALSE, message=FALSE}
# Subsetting honest_DEF in honest_DEF_nurse and honest_DEF_physician
honest_DEF_nurse <- honest_DEF[, c("typologie.f", "wageFTE_nurses", "month_wageFTE_nurses", "age.mean_nurse", "pct.female_nurse", "pct.CH_nurse", "pct.RNs_etc", "pct.LPNs", "pct.n_ass", "equipment", "AMBULANTE_KONSULTATIONEN")]
honest_DEF_physicians <- honest_DEF[, c("typologie.f", "wageFTE_physicians", "month_wageFTE_physicians","age.mean_physician", "pct.female_physician", "pct.CH_physician", "pct.chief_leading", "pct.attending_hospital", "pct.residents_students","equipment", "AMBULANTE_KONSULTATIONEN")]

# Get the list of variable names (excluding the dependent variable)
independent_vars_nurse <- names(honest_DEF_nurse)[-which(names(honest_DEF_nurse) %in% "wageFTE_nurses")]
independent_vars_physician <- names(honest_DEF_physicians)[-which(names(honest_DEF_physicians) %in% "wageFTE_physicians")]

# Create scatterplots with fitted regression lines for all variables in the nurses dataset
for (var in independent_vars_nurse) {
  # Create a plot for each variable
  plot <- ggplot(honest_DEF_nurse, aes_string(x = var, y = "wageFTE_nurses")) +
    geom_point() +
    stat_smooth(method = "lm") +
    labs(title = paste("Scatterplot with Fitted Regression Line for", var),
         x = var, y = "Wage (wageFTE_nurses)")
  
  # Print or save the plot as needed
  print(plot)
}
```


### Variables for physicians regressions

```{r scatterplots 2, echo=FALSE, warning=FALSE}


# Create scatterplots with fitted regression lines for all variables in the physician dataset
for (var in independent_vars_physician) {
  # Create a plot for each variable
  plot <- ggplot(honest_DEF_physicians, aes_string(x = var, y = "wageFTE_physicians")) +
    geom_point() +
    stat_smooth(method = "lm") +
    labs(title = paste("Scatterplot with Fitted Regression Line for", var),
         x = var, y = "Wage (wageFTE_physicians)")
  
  # Print or save the plot as needed
  print(plot)
}


```


## 4. Descriptive Data Analysis

### Descriptive Data Analysis Nurses

```{r, descriptive nurses}
honest_DEF_nurse <- honest_DEF_nurse %>%
  mutate(across(c(pct.female_nurse, pct.CH_nurse, pct.RNs_etc, pct.LPNs, pct.n_ass), ~ . * 100))

table1::table1(~ month_wageFTE_nurses + age.mean_nurse + pct.female_nurse + pct.CH_nurse + pct.RNs_etc + pct.LPNs + pct.n_ass | typologie.f, data = honest_DEF_nurse)
```

### Descriptive Data Analysis Physicians

```{r, descriptive physicians}
honest_DEF_physicians <- honest_DEF_physicians %>%
  mutate(across(c(pct.female_physician, pct.CH_physician, pct.chief_leading, pct.attending_hospital, pct.residents_students), ~ . * 100))

table1::table1(~ month_wageFTE_physicians + age.mean_physician + pct.female_physician + pct.CH_physician + pct.chief_leading + pct.attending_hospital + pct.residents_students | typologie.f, data = honest_DEF_physicians)
```



### Descriptive Data Analysis general variables

```{r, descriptive general var}
honest_DEF <- honest_DEF %>%
  mutate(AMBULANTE_KONSULTATIONEN = AMBULANTE_KONSULTATIONEN/1000)

table1::table1(~ equipment + AMBULANTE_KONSULTATIONEN | typologie.f, data = honest_DEF)
```




