---
title: "Secondary Analysis: Association Between Nurse and Physician Wages and Staffing"
author: "Aleksandra Vasic"
output: pdf_document
---
```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Packages
library(tinytex)
library(dplyr)
library(ggplot2)
library(scales)
library(patchwork)
library(extrafont)

# Set Work Directory & load merged data ----
setwd("/Volumes/research$/HONEST Team/Data/aim2_processed-data/aim2_new-merged-data/")
load("New.Merged.FTE.Data.RData")
```
\newpage
# Nurses
```{r echo=TRUE}
# Create variable "Nurse staffing level" (nurse.staff.level)
## 2020
New.Merged.FTE.2020$nurse.staff.level <- 
  New.Merged.FTE.2020$Pflege_FTE*1890/New.Merged.FTE.2020$PFLEGETAGE_STATIONAER

## 2019
New.Merged.FTE.2019$nurse.staff.level <- 
  New.Merged.FTE.2019$Pflege_FTE*1890/New.Merged.FTE.2019$PFLEGETAGE_STATIONAER

## 2018
New.Merged.FTE.2018$nurse.staff.level <- 
  New.Merged.FTE.2018$Pflege_FTE*1890/New.Merged.FTE.2018$PFLEGETAGE_STATIONAER

# Correlation Nurses - Overall
cor(New.Merged.FTE.2020$nurse.staff.level, New.Merged.FTE.2020$wageFTE_nurses, method="spearman")
cor(New.Merged.FTE.2019$nurse.staff.level, New.Merged.FTE.2019$wageFTE_nurses, method="spearman")
cor(New.Merged.FTE.2018$nurse.staff.level, New.Merged.FTE.2018$wageFTE_nurses, method="spearman")

# Correlation Nurses - Stratified by Hospital Level
## 2020
New.Merged.FTE.2020 %>%
  group_by(Hosp.Level) %>%
  summarise(correlation = cor(nurse.staff.level, wageFTE_nurses, method = "spearman"))

## 2019
New.Merged.FTE.2019 %>%
  group_by(Hosp.Level) %>%
  summarise(correlation = cor(nurse.staff.level, wageFTE_nurses, method = "spearman"))

## 2018
New.Merged.FTE.2018 %>%
  group_by(Hosp.Level) %>%
  summarise(correlation = cor(nurse.staff.level, wageFTE_nurses, method = "spearman"))


```
\newpage
## Nurse Scatterplots
### 2020 - Stratified by Hospital Level
```{r echo=TRUE}
ggplot(data = New.Merged.FTE.2020, aes(x=wageFTE_nurses, y=nurse.staff.level)) + 
  geom_point(aes(color=Hosp.Level)) +
  geom_smooth(method=lm, color="black") +
  geom_smooth(method=lm, aes(color=Hosp.Level)) +  
  labs(
    title = "Scatterplot of Wages & Nurse Staffing Levels 2020",
    x = "Wages per FTE",
    y = "Nurse Staffing Level",
    color = "Hospital Level"
  ) +
  scale_x_continuous(labels = scales::comma, limits = c(40000, 120000))
```
\newpage
### 2019 - Stratified by Hospital Level
```{r echo=TRUE}
ggplot(data = New.Merged.FTE.2019, aes(x=wageFTE_nurses, y=nurse.staff.level)) + 
  geom_point(aes(color=Hosp.Level)) +
  geom_smooth(method=lm, color="black") +
  geom_smooth(method=lm, aes(color=Hosp.Level)) +  
  labs(
    title = "Scatterplot of Wages & Nurse Staffing Levels 2019",
    x = "Wages per FTE",
    y = "Nurse Staffing Level",
    color = "Hospital Level"
  ) +
  scale_x_continuous(labels = scales::comma, limits = c(40000, 125000))
```
\newpage
### 2018 - Stratified by Hospital Level
```{r echo=TRUE}
ggplot(data = New.Merged.FTE.2018, aes(x=wageFTE_nurses, y=nurse.staff.level)) + 
  geom_point(aes(color=Hosp.Level)) +
  geom_smooth(method=lm, color="black") +
  geom_smooth(method=lm, aes(color=Hosp.Level)) +  
  labs(
    title = "Scatterplot of Wages & Nurse Staffing Levels 2018",
    x = "Wages per FTE",
    y = "Nurse Staffing Level",
    color = "Hospital Level"
  ) +
  scale_x_continuous(labels = scales::comma, limits = c(35000, 150000))
```

\newpage
# Physicians
```{r echo=TRUE}
# Create variable "Physician staffing level" (phys.staff.level)
## 2020
New.Merged.FTE.2020$phys.staff.level <- 
  New.Merged.FTE.2020$Arzt_FTE*2475/New.Merged.FTE.2020$PFLEGETAGE_STATIONAER

## 2019
New.Merged.FTE.2019$phys.staff.level <- 
  New.Merged.FTE.2019$Arzt_FTE*2475/New.Merged.FTE.2019$PFLEGETAGE_STATIONAER

## 2018
New.Merged.FTE.2018$phys.staff.level <- 
  New.Merged.FTE.2018$Arzt_FTE*2475/New.Merged.FTE.2018$PFLEGETAGE_STATIONAER

# Correlation Physicians - Overall
cor(New.Merged.FTE.2020$phys.staff.level, New.Merged.FTE.2020$wageFTE_physicians, method="spearman")
cor(New.Merged.FTE.2019$phys.staff.level, New.Merged.FTE.2019$wageFTE_physicians, method="spearman")
cor(New.Merged.FTE.2018$phys.staff.level, New.Merged.FTE.2018$wageFTE_physicians, method="spearman")

# Correlation Physicians - Stratified by Hospital Level
## 2020
New.Merged.FTE.2020 %>%
  group_by(Hosp.Level) %>%
  summarise(correlation = cor(phys.staff.level, wageFTE_physicians, method = "spearman"))

## 2019
New.Merged.FTE.2019 %>%
  group_by(Hosp.Level) %>%
  summarise(correlation = cor(phys.staff.level, wageFTE_physicians, method = "spearman"))

## 2018
New.Merged.FTE.2018 %>%
  group_by(Hosp.Level) %>%
  summarise(correlation = cor(phys.staff.level, wageFTE_physicians, method = "spearman"))
```
\newpage
## Physician Scatterplots
### 2020 - Stratified by Hospital Level
\newpage
```{r echo=TRUE}
ggplot(data = New.Merged.FTE.2020, aes(x=wageFTE_physicians, y=phys.staff.level)) + 
  geom_point(aes(color=Hosp.Level)) +
  geom_smooth(method=lm, color="black") +
  geom_smooth(method=lm, aes(color=Hosp.Level)) +  
  labs(
    title = "Scatterplot of Wages & Physician Staffing Levels 2020",
    x = "Wages per FTE",
    y = "Physician Staffing Level",
    color = "Hospital Level"
  ) +
  scale_x_continuous(labels = scales::comma, limits = c(80000, 400000))
```
\newpage
## 2019 - Stratified by Hospital Level
```{r echo=TRUE}
ggplot(data = New.Merged.FTE.2019, aes(x=wageFTE_physicians, y=phys.staff.level)) + 
  geom_point(aes(color=Hosp.Level)) +
  geom_smooth(method=lm, color="black") +
  geom_smooth(method=lm, aes(color=Hosp.Level)) +  
  labs(
    title = "Scatterplot of Wages & Physician Staffing Levels 2019",
    x = "Wages per FTE",
    y = "Physician Staffing Level",
    color = "Hospital Level"
  ) +
  scale_x_continuous(labels = scales::comma, limits = c(100000, 400000))
```
\newpage
### 2018 - Stratified by Hospital Level
```{r echo=TRUE}
ggplot(data = New.Merged.FTE.2018, aes(x=wageFTE_physicians, y=phys.staff.level)) + 
  geom_point(aes(color=Hosp.Level)) +
  geom_smooth(method=lm, color="black") +
  geom_smooth(method=lm, aes(color=Hosp.Level)) +  
  labs(
    title = "Scatterplot of Wages & Physician Staffing Levels 2018",
    x = "Wages per FTE",
    y = "Physician Staffing Level",
    color = "Hospital Level"
  ) +
  scale_x_continuous(labels = scales::comma, limits = c(100000, 400000))
```

\newpage
# Figures for Publication
## 2020
```{r echo=TRUE}
font_import(pattern = "Times", prompt = FALSE)
loadfonts(device = "pdf")

nurse20 <- ggplot(data = New.Merged.FTE.2020, aes(x=wageFTE_nurses, y=nurse.staff.level)) + 
  geom_point(aes(color=Hosp.Level)) +
  geom_smooth(method=lm, color="black") +
  geom_smooth(method=lm, aes(color=Hosp.Level)) +  
  labs(title = "(A)",
       x = "Wages per FTE",
       y = "Staffing Level",
       colour = "Hospital Level") +
  theme_minimal(base_family = "Times New Roman") +
  scale_x_continuous(labels = scales::comma, limits = c(35000, 150000))

phys20 <- ggplot(data = New.Merged.FTE.2020, aes(x=wageFTE_physicians, y=phys.staff.level)) + 
  geom_point(aes(color=Hosp.Level)) +
  geom_smooth(method=lm, color="black") +
  geom_smooth(method=lm, aes(color=Hosp.Level)) +  
  labs(title = "(B)",
       x = "Wages per FTE",
       y = "Staffing Level",
       colour = "Hospital Level") +
  theme_minimal(base_family = "Times New Roman") +
  scale_x_continuous(labels = scales::comma, limits = c(100000, 400000))


combined_20 <- nurse20 + phys20 +
  plot_annotation(
    title   = "Scatterplot of Wages &  Staffing Levels, 2020",
    caption = "(A) Nurses, (B) Physicians",
    theme   = theme(
      plot.title = element_text(
        family = "Times New Roman",
        size   = 16,
        face   = "bold",
        hjust  = 0.5
      ),
      plot.caption = element_text(
        family = "Times New Roman",
        size   = 10,
        hjust  = 0,
        margin = margin(t = 10)
      )
    )
  ) &
  theme(
    text = element_text(family = "Times New Roman")
  )

print(combined_20)
```

\newpage
## 2019
```{r echo=TRUE}
font_import(pattern = "Times", prompt = FALSE)
loadfonts(device = "pdf")

nurse19 <- ggplot(data = New.Merged.FTE.2019, aes(x=wageFTE_nurses, y=nurse.staff.level)) + 
  geom_point(aes(color=Hosp.Level)) +
  geom_smooth(method=lm, color="black") +
  geom_smooth(method=lm, aes(color=Hosp.Level)) +  
  labs(title = "(A)",
       x = "Wages per FTE",
       y = "Staffing Level",
       colour = "Hospital Level") +
  theme_minimal(base_family = "Times New Roman") +
  scale_x_continuous(labels = scales::comma, limits = c(35000, 150000))

phys19 <- ggplot(data = New.Merged.FTE.2019, aes(x=wageFTE_physicians, y=phys.staff.level)) + 
  geom_point(aes(color=Hosp.Level)) +
  geom_smooth(method=lm, color="black") +
  geom_smooth(method=lm, aes(color=Hosp.Level)) +  
  labs(title = "(B)",
       x = "Wages per FTE",
       y = "Staffing Level",
       colour = "Hospital Level") +
  theme_minimal(base_family = "Times New Roman") +
  scale_x_continuous(labels = scales::comma, limits = c(100000, 400000))


combined_19 <- nurse19 + phys19 +
  plot_annotation(
    title   = "Scatterplot of Wages &  Staffing Levels, 2019",
    caption = "(A) Nurses, (B) Physicians",
    theme   = theme(
      plot.title = element_text(
        family = "Times New Roman",
        size   = 16,
        face   = "bold",
        hjust  = 0.5
      ),
      plot.caption = element_text(
        family = "Times New Roman",
        size   = 10,
        hjust  = 0,
        margin = margin(t = 10)
      )
    )
  ) &
  theme(
    text = element_text(family = "Times New Roman")
  )

print(combined_19)
```

\newpage
## 2018
```{r echo=TRUE}
font_import(pattern = "Times", prompt = FALSE)
loadfonts(device = "pdf")

nurse18 <- ggplot(data = New.Merged.FTE.2018, aes(x=wageFTE_nurses, y=nurse.staff.level)) + 
  geom_point(aes(color=Hosp.Level)) +
  geom_smooth(method=lm, color="black") +
  geom_smooth(method=lm, aes(color=Hosp.Level)) +  
  labs(title = "(A)",
       x = "Wages per FTE",
       y = "Staffing Level",
       colour = "Hospital Level") +
  theme_minimal(base_family = "Times New Roman") +
  scale_x_continuous(labels = scales::comma, limits = c(35000, 150000))

phys18 <- ggplot(data = New.Merged.FTE.2018, aes(x=wageFTE_physicians, y=phys.staff.level)) + 
  geom_point(aes(color=Hosp.Level)) +
  geom_smooth(method=lm, color="black") +
  geom_smooth(method=lm, aes(color=Hosp.Level)) +  
  labs(title = "(B)",
       x = "Wages per FTE",
       y = "Staffing Level",
       colour = "Hospital Level") +
  theme_minimal(base_family = "Times New Roman") +
  scale_x_continuous(labels = scales::comma, limits = c(100000, 400000))


combined_18 <- nurse18 + phys18 +
  plot_annotation(
    title   = "Scatterplot of Wages &  Staffing Levels, 2018",
    caption = "(A) Nurses, (B) Physicians",
    theme   = theme(
      plot.title = element_text(
        family = "Times New Roman",
        size   = 16,
        face   = "bold",
        hjust  = 0.5
      ),
      plot.caption = element_text(
        family = "Times New Roman",
        size   = 10,
        hjust  = 0,
        margin = margin(t = 10)
      )
    )
  ) &
  theme(
    text = element_text(family = "Times New Roman")
  )

print(combined_18)
```

